The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2014-09-19 paper/mri_paper.tex
fixed appendix.
more appendix

We conduct a formal multiscale analysis. Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on fast and slow scales. The relative scalings of the fast and slow scales are determined as follows. When our four equations are linearized for axisymmetric perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$. The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation. The scalings are

@@

The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix \ref{app:appendix_nonlinear}. $\mathbf{N}$ becomes

@@

Equation \ref{ampeq1} is an amplitude equation -- a function of the slow variables $Z$ and $T$. We solve this initial value problem using Dedalus to obtain the asymptotic behavior of $\alpha$. This is the saturation amplitude.

\section{Results}

@@

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.5]{allorders_velocity.eps}
\caption{First order (left) and second order (right) velocity perturbations. Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds. The colorbar represents azimuthal velocity.}
\end{figure*}

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.5]{allorders_Bfield.eps}
\caption{First order (left) and second order (right) velocity perturbations. Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds. The colorbar represents azimuthal velocity.}
\end{figure*}

@@

\newpage
\begin{subappendices}
\begin{widetext}
\section*{Appendix A. Nonlinear term expansion}\label{app:appendix_nonlinear}

\beq
\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
\eeq

@@

\beq
N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1)
\eeq

@@

\end{widetext}
\end{subappendices}

>>>>

We conduct a formal multiscale analysis. Our perturbations are characterized in terms of fast and slow-moving variables, in order to simultaneously track their evolution on fast and slow scales. The relative scalings of the fast and slow scales are determined as follows. When our four equations are linearized for axisymmetric perturbations of the form $e^{\sigma t + i k_x x + i k_z z}$, we can derive the linear dispersion relation, which is fourth order in $\sigma$ (Appendix \ref{app:dispersion}). The fast and slow variables are then defined such that each of the temporal and spatial eigenvalues appear at the same lowest order in the linear dispersion relation. The scalings are

@@

The multiscale expansion of the operators and the perturbation series expansion of the fluid quantities combine to make the expansion of the nonlinear vector $\mathbf{N}$ expansive, and we relegate the expansion to Appendix \ref{app:nonlinear}. $\mathbf{N}$ becomes

@@

Equation \ref{ampeq1} is an amplitude equation -- a function of the slow variables $Z$ and $T$. We solve this initial value problem using Dedalus to obtain the asymptotic behavior of $\alpha$. This is the saturation amplitude. We use the saturation amplitude as $\alpha$ and the radial component of computed on the Chebyshev grid to plot $\mathbf{V}$. We show both first order ($\epsilon\mathbf{V}_1$), second order ($\epsilon^2\mathbf{V}_2$), and combined first and second order fluid perturbations ($\epsilon\mathbf{V}_1$ + ($\epsilon^2\mathbf{V}_2$). Figure \ref{allorders_velocity} shows the velocity perturbations, and Figure \ref{allorders_Bfield} shows the magnetic field perturbations.

\section{Parameter Space}

@@

\begin{figure*}[h!]
\centering
\includegraphics[scale=0.5]{allorders_velocity.eps}
\caption{First order (left) and second order (right) velocity perturbations. Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds. The colorbar represents azimuthal velocity.}\label{allorders_velocity}
\end{figure*}

\begin{figure*}[h!]\label{allorders_Bfield}
\centering
\includegraphics[scale=0.5]{allorders_Bfield.eps}
\caption{First order (left) and second order (right) velocity perturbations. Streamlines represent velocity in the vertical-radial plane, where thicker streamlines correspond to faster speeds. The colorbar represents azimuthal velocity.}
\end{figure*}

@@

\clearpage
\appendix

\section{A. Linear dispersion relation}\label{app:dispersion}
%Note: I don't expect this to go in the actual paper, but it could be useful in the meeting.

The linear dispersion relation, which determines the variable scalings in the multiscale analysis. \\

\beq
\begin{split}
& \frac{B_{0}^{4} k_{x}^{2} k_{z}^{4}}{16 \pi^{2}} + \frac{B_{0}^{4} k_{z}^{6}}{16 \pi^{2}} - \frac{B_{0}^{2} \Omega_{0} k_{z}^{4} q}{2 \pi} - 2 \Omega_{0} k_{z}^{2} q \sigma^{2} - \frac{4 \sigma}{\reym} \Omega_{0} k_{x}^{2} k_{z}^{2} q - \frac{4 \sigma}{\reym} \Omega_{0} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{x}^{4} k_{z}^{2} q - \frac{4 \Omega_{0}}{\reym^{2}} k_{x}^{2} k_{z}^{4} q - \frac{2 \Omega_{0}}{\reym^{2}} k_{z}^{6} q \\
& - k_{x}^{2} \sigma^{4} - k_{z}^{2} \sigma^{4} + 4 k_{z}^{2} \sigma^{2} - \frac{2 \sigma^{3}}{\reym} k_{x}^{4} - \frac{4 \sigma^{3}}{\reym} k_{x}^{2} k_{z}^{2} + \frac{8 \sigma}{\reym} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reym} k_{z}^{4} + \frac{8 \sigma}{\reym} k_{z}^{4} - \frac{k_{x}^{6} \sigma^{2}}{\reym^{2}} - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{4} k_{z}^{2} + \frac{4 k_{x}^{4}}{\reym^{2}} k_{z}^{2} \\
& - \frac{3 \sigma^{2}}{\reym^{2}} k_{x}^{2} k_{z}^{4} + \frac{8 k_{x}^{2}}{\reym^{2}} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reym^{2}} + \frac{4 k_{z}^{6}}{\reym^{2}} - \frac{2 \sigma^{3}}{\reye} k_{x}^{4} - \frac{4 \sigma^{3}}{\reye} k_{x}^{2} k_{z}^{2} - \frac{2 \sigma^{3}}{\reye} k_{z}^{4} - \frac{4 k_{x}^{6} \sigma^{2}}{\reye \reym} -  \frac{12 k_{x}^{4} k_{z}^{2} \sigma^{2}}{\reye \reym} - \frac{12 k_{x}^{2} k_{z}^{4} \sigma^{2}}{\reye \reym} \\
& - \frac{4 k_{z}^{6} \sigma^{2}}{\reye \reym} - \frac{2 k_{x}^{8} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye \reym^{2}} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye \reym^{2}} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye \reym^{2}} - \frac{2 k_{z}^{8} \sigma}{\reye \reym^{2}} - \frac{k_{x}^{6} \sigma^{2}}{\reye^{2}} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{4} k_{z}^{2} - \frac{3 \sigma^{2}}{\reye^{2}} k_{x}^{2} k_{z}^{4} - \frac{k_{z}^{6} \sigma^{2}}{\reye^{2}} \\
& - \frac{2 k_{x}^{8} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{6} k_{z}^{2} \sigma}{\reye^{2} \reym} - \frac{12 k_{x}^{4} k_{z}^{4} \sigma}{\reye^{2} \reym} - \frac{8 k_{x}^{2} k_{z}^{6} \sigma}{\reye^{2} \reym} - \frac{2 k_{z}^{8} \sigma}{\reye^{2} \reym} - \frac{k_{x}^{10}}{\reye^{2} \reym^{2}} - \frac{5 k_{x}^{8} k_{z}^{2}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{6} k_{z}^{4}}{\reye^{2} \reym^{2}} - \frac{10 k_{x}^{4} k_{z}^{6}}{\reye^{2} \reym^{2}} \\
& - \frac{5 k_{x}^{2} k_{z}^{8}}{\reye^{2} \reym^{2}} - \frac{k_{z}^{10}}{\reye^{2} \reym^{2}} = 0
\end{split}
\eeq

\section{B. Expansion of Nonlinear Terms}\label{app:nonlinear}
\beq
\mathbf{N} = \epsilon^2\mathbf{N_2} \, + \, \epsilon^3\mathbf{N_3} \, + \, \mathcal{O}(\epsilon^4)
\eeq

@@

\beq
N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1).\\
\eeq

@@
